### 2020-12-30 Paper/AAC_Pseudocode/Advantage_Actor_Critic_Alg.tex
Tlatex file and script to save algorithm into a single pdf file and then transform it into pdf

%\documentclass[preview]{scrartcl}
% set margin
%\usepackage[a4paper, total={6in, 8in}]{geometry}
%\documentclass[varwidth]{standalone}
%\usepackage[utf8]{inputenc}
%\usepackage[subpreambles]{standalone}

\documentclass{article}
\usepackage{blindtext}
%\usepackage[paperheight=5in,paperwidth=8.25in,margin=0.1in,heightrounded,showframe]{geometry}
\usepackage[paperheight=3in,paperwidth=5in,margin=0.05in]{geometry}

%Remove page number
\pagenumbering{gobble}

\usepackage{amsfonts}
\usepackage{algorithm2e}
%\usepackage{graphicx}
\usepackage{amsmath}

\begin{document}
%\include{pseudocode}

\begin{algorithm}
 \textbf{Input}: differentiable policy and state-value function parameterization $\pi(a|s;\boldsymbol{\theta})$ and $V(s;|s;\boldsymbol{w})$\;
 \textbf{Parameters}: step sizes $\alpha^{\boldsymbol{\theta}}>0,\  \alpha^{\boldsymbol{w}}>0$ and discount factor $\gamma \in [0,1]$\;
 Initialize policy and state-value parameters $\boldsymbol{\theta} \in \mathbb{R}^d$ and $\boldsymbol{w} \in \mathbb{R}^{d'}$\;
 \For{Episode = 1, 2, ...}{
  \While{$S_t$ is not terminal (for each time step $t$)}{
    Sample action $A_t \sim \pi(\cdot|S_t;\boldsymbol{\theta})$ \;
    Take action $A_t$ and observe $S_{t+1}, R_{t+1}$\;
    \If{$S_{t+1}$ is terminal}{
      $V(S_{t+1}; \boldsymbol{w}) = 0$\;
    }
    $\delta \gets R_{t+1} + \gamma V(S_{t+1}; \boldsymbol{w}) - V(S_{t}; \boldsymbol{w})$\;
    $\boldsymbol{w} \gets \boldsymbol{w} - \alpha^{\boldsymbol{w}} \delta \nabla V(S_{t}; \boldsymbol{w})$\;
    $\boldsymbol{\theta} \gets \boldsymbol{\theta} + \alpha^{\boldsymbol{\theta}} \delta \nabla \log{\pi(A_{t}|S_{t}; \boldsymbol{\theta})}$\;
    $S_t \gets S_{t+1}$\;
  }
 }
\end{algorithm}

\end{document}